---
title: "Unit Root Tests"
author: "J Coetsee - 19491050"
date: "31/05/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r}
library(tidyverse)
library(scales)
theme_set(theme_light())
library(lubridate)
library(panelvar)
library(gridExtra)
library(plm)
library(tseries)
library(huxtable)


#sourcing in functions from code folder 

source("code/growth_fun.R")

data <- read.csv("C:/Projects/Time_Series/PVAR_19491050/data/joined.csv")
```

# Series Demeaning:

```{r}

#creating cols of demeaned data:
data <- data %>% 
    group_by(State) %>% 
    mutate(Gini_dem = (Gini - mean(Gini)),
           Inc_dem = (realpc_income - mean(realpc_income))) 

### creating Gini data for levels

Gini_demeaned <- data %>% 
  select(c(Year, State, Gini_dem))

#this is the dataset used for the tests:

gini_data_wide <-  data.frame(split(Gini_demeaned$Gini_dem, Gini_demeaned$State))

### Creating Income Data for Levels

inc_demeaned <- data %>% 
  select(c(Year, State, Inc_dem))

#and the df used for tests:

inc_data_wide <-  data.frame(split(inc_demeaned$Inc_dem, inc_demeaned$State))

```


# using purtest: 

```{r}
#example:
#data("Grunfeld", package = "plm")
#y <- data.frame(split(Grunfeld$inv, Grunfeld$firm)) # individuals in columns
#purtest(y, pmax = 4, exo = "intercept", test = "madwu")
## same via formula interface
#purtest(inv ~ 1, data = Grunfeld, index = c("firm", "year"), pmax = 4, test = "madwu")
```

### Unit Root Tests, the level of Gini. Four tests, "levinlin", "ips", "madwu" and "hadri". 


naming convention:

gtest1l - where 'g' = gini, 'test1' = intercept and 'l' = levinlin

e.g. incgtest2a = income growth, trend and intercept, using madwu (ADF)

```{r}

# LLC:

# Gini level - intercept

gtest1l <- purtest(gini_data_wide,
        exo = "intercept",
        test = "levinlin",
        lags = "AIC",
        pmax = 10)

summary(gtest1l)
print(gtest1l)

# Gini level - intercept + Trend

gtest2l <- purtest(gini_data_wide,
        exo = "trend",
        test = "levinlin",
        lags = "AIC",
        pmax = 10)

summary(gtest2l)
print(gtest2l)

# IPS:

# Gini level - intercept

gtest1i <- purtest(gini_data_wide,
        exo = "intercept",
        test = "ips",
        lags = "AIC",
        pmax = 10)

summary(gtest1i)
print(gtest1i)

# Gini level - intercept + Trend

gtest2i <- purtest(gini_data_wide,
        exo = "trend",
        test = "ips",
        lags = "AIC",
        pmax = 10)

summary(gtest2i)
print(gtest2i)

# ADF:

# Gini level - intercept

gtest1a <- purtest(gini_data_wide,
        exo = "intercept",
        test = "madwu",
        lags = "AIC",
        pmax = 10)

summary(gtest1a)
print(gtest1a)

# Gini level - intercept + Trend

gtest2a <- purtest(gini_data_wide,
        exo = "trend",
        test = "madwu",
        lags = "AIC",
        pmax = 10)

summary(gtest2a)
print(gtest2a)


# hadri

# Gini level - intercept

gtest1h <- purtest(gini_data_wide,
        exo = "intercept",
        test = "hadri",
        lags = "AIC",
        pmax = 10)

summary(gtest1h)
print(gtest1h)

# Gini level - intercept + Trend

gtest2h <- purtest(gini_data_wide,
        exo = "trend",
        test = "hadri",
        lags = "AIC",
        pmax = 10)

summary(gtest2h)
print(gtest2h)

```

Unit Root Tests, the *growth* of Gini. Four tests, "levinlin", "ips", "madwu" and "hadri"

```{r}

ginigrowth_demeaned <- Gini_demeaned %>% 
  group_by(State) %>% 
  mutate(growth = c(NA,diff(Gini_dem))/lag(Gini_dem, 1)) %>% slice(-1)

ginigrowth_data_wide <- data.frame(split(ginigrowth_demeaned$growth, ginigrowth_demeaned$State))

### LLC:

# Gini level - intercept

ggtest1l <- purtest(ginigrowth_data_wide,
        exo = "intercept",
        test = "levinlin",
        lags = "AIC",
        pmax = 10)

summary(ggtest1l)
print(ggtest1l)

# Gini level - intercept + Trend

ggtest2l <- purtest(ginigrowth_data_wide,
        exo = "trend",
        test = "levinlin",
        lags = "AIC",
        pmax = 10)

summary(ggtest2l)
print(ggtest2l)

# IPS:

# Gini level - intercept

ggtest1i <- purtest(ginigrowth_data_wide,
        exo = "intercept",
        test = "ips",
        lags = "AIC",
        pmax = 10)

summary(ggtest1i)
print(ggtest1i)

# Gini level - intercept + Trend

ggtest2i <- purtest(ginigrowth_data_wide,
        exo = "trend",
        test = "ips",
        lags = "AIC",
        pmax = 10)

summary(ggtest2i)
print(ggtest2i)

# ADF:

# Gini level - intercept

ggtest1a <- purtest(ginigrowth_data_wide,
        exo = "intercept",
        test = "madwu",
        lags = "AIC",
        pmax = 10)

summary(ggtest1a)
print(ggtest1a)

# Gini level - intercept + Trend

ggtest2a <- purtest(ginigrowth_data_wide,
        exo = "trend",
        test = "madwu",
        lags = "AIC",
        pmax = 10)

summary(ggtest2a)
print(ggtest2a)


# hadri

# Gini level - intercept

ggtest1h <- purtest(ginigrowth_data_wide,
        exo = "intercept",
        test = "hadri",
        lags = "AIC",
        pmax = 10)

summary(ggtest1h)
print(ggtest1h)

# Gini level - intercept + Trend

ggtest2h <- purtest(ginigrowth_data_wide,
        exo = "trend",
        test = "hadri",
        lags = "AIC",
        pmax = 10)

summary(ggtest2h)
print(ggtest2h)


```

####### Income Unit Root Tests


Unit Root Tests, the *level* of Income using "inc_data_wide". Four tests, "levinlin", "ips", "madwu" and "hadri"

```{r}

# LLC:

# Gini level - intercept

inctest1l <- purtest(inc_data_wide,
        exo = "intercept",
        test = "levinlin",
        lags = "AIC",
        pmax = 10)

summary(inctest1l)
print(inctest1l)

# Gini level - intercept + Trend

inctest2l <- purtest(inc_data_wide,
        exo = "trend",
        test = "levinlin",
        lags = "AIC",
        pmax = 10)

summary(inctest2l)
print(inctest2l)

# IPS:

# Gini level - intercept

inctest1i <- purtest(inc_data_wide,
        exo = "intercept",
        test = "ips",
        lags = "AIC",
        pmax = 10)

summary(inctest1i)
print(inctest1i)

# Gini level - intercept + Trend

inctest2i <- purtest(inc_data_wide,
        exo = "trend",
        test = "ips",
        lags = "AIC",
        pmax = 10)

summary(inctest2i)
print(inctest2i)

# ADF:

# Gini level - intercept

inctest1a <- purtest(inc_data_wide,
        exo = "intercept",
        test = "madwu",
        lags = "AIC",
        pmax = 10)

summary(inctest1a)
print(inctest1a)

# Gini level - intercept + Trend

inctest2a <- purtest(inc_data_wide,
        exo = "trend",
        test = "madwu",
        lags = "AIC",
        pmax = 10)

summary(inctest2a)
print(inctest2a)


# hadri

# Gini level - intercept

inctest1h <- purtest(inc_data_wide,
        exo = "intercept",
        test = "hadri",
        lags = "AIC",
        pmax = 10)

summary(inctest1h)
print(inctest1h)

# Gini level - intercept + Trend

inctest2h <- purtest(inc_data_wide,
        exo = "trend",
        test = "hadri",
        lags = "AIC",
        pmax = 10)

summary(inctest2h)
print(inctest2h)

```

Unit Root Tests, the *growth* of Income. Four tests, "levinlin", "ips", "madwu" and "hadri"

```{r}

#growth data for income:

incgrowth_demeaned <- inc_demeaned %>% 
  group_by(State) %>% 
  mutate(growth = c(NA,diff(Inc_dem))/lag(Inc_dem, 1)) %>% slice(-1)

incgrowth_data_wide <- data.frame(split(incgrowth_demeaned$growth, incgrowth_demeaned$State))

### LLC:

# Gini level - intercept

incgtest1l <- purtest(incgrowth_data_wide,
        exo = "intercept",
        test = "levinlin",
        lags = "AIC",
        pmax = 10)

summary(incgtest1l)
print(incgtest1l)

# Gini level - intercept + Trend

incgtest2l <- purtest(incgrowth_data_wide,
        exo = "trend",
        test = "levinlin",
        lags = "AIC",
        pmax = 10)

summary(incgtest2l)
print(incgtest2l)

# IPS:

# Gini level - intercept

incgtest1i <- purtest(incgrowth_data_wide,
        exo = "intercept",
        test = "ips",
        lags = "AIC",
        pmax = 10)

summary(incgtest1i)
print(incgtest1i)

# Gini level - intercept + Trend

incgtest2i <- purtest(incgrowth_data_wide,
        exo = "trend",
        test = "ips",
        lags = "AIC",
        pmax = 10)

summary(incgtest2i)
print(incgtest2i)

# ADF:

# Gini level - intercept

incgtest1a <- purtest(incgrowth_data_wide,
        exo = "intercept",
        test = "madwu",
        lags = "AIC",
        pmax = 10)

summary(incgtest1a)
print(incgtest1a)

# Gini level - intercept + Trend

incgtest2a <- purtest(incgrowth_data_wide,
        exo = "trend",
        test = "madwu",
        lags = "AIC",
        pmax = 10)

summary(incgtest2a)
print(incgtest2a)


# hadri

# Gini level - intercept

incgtest1h <- purtest(incgrowth_data_wide,
        exo = "intercept",
        test = "hadri",
        lags = "AIC",
        pmax = 10)

summary(incgtest1h)
print(incgtest1h)

# Gini level - intercept + Trend

incgtest2h <- purtest(incgrowth_data_wide,
        exo = "trend",
        test = "hadri",
        lags = "AIC",
        pmax = 10)

summary(incgtest2h)
print(incgtest2h)


```

The full results are thus:

```{r}

# Gini Levels

#LLC:
gtest1l$statistic
gtest2l$statistic
# IPS:
gtest1i$statistic
gtest2i$statistic
# ADF:
gtest1a$statistic
gtest2a$statistic
# Hadri:
gtest1h$statistic
gtest2h$statistic

# Gini growth

#LLC:
ggtest1l$statistic
ggtest2l$statistic
# IPS:
ggtest1i$statistic
ggtest2i$statistic
# ADF:
ggtest1a$statistic
ggtest2a$statistic
# Hadri:
ggtest1h$statistic
ggtest2h$statistic

# Income Levels

#LLC:
inctest1l$statistic
inctest2l$statistic
# IPS:
inctest1i$statistic
inctest2i$statistic
# ADF:
inctest1a$statistic
inctest2a$statistic
# Hadri:
inctest1h$statistic
inctest2h$statistic

# Income growth

#LLC:
incgtest1l$statistic
incgtest2l$statistic
# IPS:
incgtest1i$statistic
incgtest2i$statistic
# ADF:
incgtest1a$statistic
incgtest2a$statistic
# Hadri:
incgtest1h$statistic
incgtest2h$statistic

```

### Table of Results: 

This was not completed, too much difficulty trying to make a table 

```{r}

#create list

unitlist <- list(gtest1l,gtest2l,gtest1i,gtest2i,gtest1a,gtest2a,gtest1h,gtest2h,ggtest1l,ggtest2l,ggtest1i,ggtest2i,ggtest1a,ggtest2a,ggtest1h,ggtest2h,inctest1l,inctest2l,inctest1i,inctest2i,inctest1a,inctest2a,inctest1h,inctest2h,incgtest1l,incgtest2l,incgtest1i,incgtest2i,incgtest1a,incgtest2a,incgtest1h,incgtest2h)

```

