---
title: "Data"
author: "J Coetsee"
date: "13/04/2021"
output: html_document
---

# Packages and code:

```{r}

library(tidyverse)
library(scales)
theme_set(theme_light())
library(lubridate)
library(panelvar)
library(gridExtra)

#sourcing in functions from code folder 

source("code/growth_fun.R")

```

# Read in the Data: 

Income

```{r}

income <- read.csv("data/SAINC1__ALL_AREAS_1929_2020.csv")

```

Inequality dataset:

```{r}
inequality <- readxl::read_xls("data/Frank_Gini_2018.xls")

#dropping non-congruous states (those that weren't part of the US since 1929, and filtering for years post 1929)
inequality <-inequality %>% 
  filter(State != "United States",
         State != "Alaska", 
         State != "Hawaii") %>% 
  filter(Year >= 1929) 

# The dropped Rows are hawaii, Alaska and United States

```

CPI Data

```{r}

CPI <- readxl::read_xlsx("data/CPI_US.xlsx") %>% 
  rename(Year = "CPI for All Urban Consumers (CPI-U)",
         Index = ...2) %>% 
  slice(-c(1:11)) %>% 
  mutate(Year = as.numeric(Year),
         Index = as.numeric(Index))

```


# Make list of Contiguous US states - statenames

```{r}
statenames <- ineq1930 %>%
  select(State) %>% 
  unique()
```

############ Inequality Wrangling ##############


# Inequality US as a whole:

```{r}

#For US as a whole:

US_ineq <- inequality %>% 
  filter(State == "United States") %>% 
  filter(Year >= 1929)

```

# ineq1930: replacing coeffs with growth rates year on year, and calling this dataset ineq1930: 

```{r}
ineq1930 <- inequality %>% 
  group_by(State) %>% 
  mutate(across(Atkin05:Theil, growth_fun)) %>% 
  filter(Year != "1929")

```

############### Income Set wrangling ##############


# Tidy Income data:

```{r}
#Making the dataset long and deselecting irrelevant variables

income <- income %>% 
  gather(key = Year, value = value, X1929:X2020) %>% 
  select(-c(TableName, LineCode, IndustryClassification)) %>%
  mutate(Year = gsub("X", "", Year))

# make pcincome df which only focuses on per capita income by state

pcincome <- income %>% 
  filter(Description == "Per capita personal income (dollars) 2/") %>% 
  rename(pc_income = value,
         State = GeoName) %>% 
  select(-c("Description", "Unit")) %>% 
  mutate(Year = as.numeric(Year),
         pc_income = as.numeric(pc_income)) %>% 
  filter(State %in% statenames$State) 

# income for US as a whole:

USincome <- income %>% 
  filter(Description == "Per capita personal income (dollars) 2/") %>% 
  rename(pc_income = value,
         State = GeoName) %>% 
  select(-c("Region", "Description", "Unit")) %>% 
  mutate(Year = as.numeric(Year),
         pc_income = as.numeric(pc_income)) %>% 
  filter(State == "United States")

USincome_defl <-  left_join(USincome, CPI, by = ("Year" = "Year")) %>% 
  mutate(realpc_income = pc_income/Index)

```

Deflating pc income

```{r}

pcdeflated <- left_join(pcincome, CPI, by = ("Year" = "Year")) %>% 
  mutate(realpc_income = pc_income/Index)

```


# pc_joined = full joined dataset of ineq1930 and pcincome

```{r}

pc_joined <- left_join(ineq1930, pcincome, by = c("State" = "State", "Year" = "Year")) 

```

# Descriptive Statistics Inequality:

```{r}
#FIGURE 1: 

# this is an exact replica of  graph a in the paper, wrt lnrealpc_income:

p1 <- USincome_defl %>% 
  mutate(realpc_income = pc_income/(Index/100),
         lnrealpc_income = log(realpc_income)) %>% 
  filter(Year >= 1930,
         Year <= 2018) %>% 
  ggplot()+
  geom_line(aes(Year, lnrealpc_income),  colour = "red") +
  labs(y = "Log of Real Income",
       title = "A")

# this is an exact replica of graph b in the paper, wrt growth in real pc income:  
  
p2 <- USincome_defl %>% 
      filter(Year >= 1930,
         Year <= 2018) %>% 
  mutate(realpc_income = pc_income/(Index/100),
         pc_incomegrowth = (realpc_income - lag(realpc_income))/lag(realpc_income)) %>% 
  ggplot() +
  geom_line(aes(Year, pc_incomegrowth), colour = "steelblue") +
  labs(y = "Income Growth",
       title = "B")

#US income plots:

inc_plot <- grid.arrange(p1, p2, nrow = 1,
             top = "Figure 1",
             bottom = "A) United States average of real income per capita and B) real income per capita growth at the state level, 1930–2018. Author's own calculations.")


```

US inequality graphs:

```{r}

#FIGURE 2: 

# Gini

US_ineq <- inequality %>% 
  filter(Year >= 1930) %>% 
  gather(key = Measure, value = Value, Atkin05:Theil)

US_ineq_facet <-  US_ineq %>% 
  group_by(Year, Measure) %>% 
  mutate(mean_measure = mean(Value)) %>% 
  ggplot()+
  geom_line(aes(Year, mean_measure, colour = Measure), size = 1) +
  facet_wrap(~ Measure, scales = "free") +
  labs(title = "Figure 2: US State-level Average of Inequality Measures",
       caption = "1930-2018. Author's own calculations.",
       y = "") +
  theme(plot.caption.position = "plot",
        strip.background = element_blank(),
        strip.text.x = element_blank())

# by top 15 states alphabetically: to repeat each geom_line for the US for each facet, use transform to make the State variable NULL - I struggled with this for a long time

inequality %>% 
  filter(State == head(State, n = 16)) %>% 
  ggplot() +
  geom_line(aes(Year, Gini, colour = State)) +
  geom_line(data = transform(US_ineq, State = NULL), aes(Year, Gini), colour = "Black", alpha = 0.5) +
  facet_wrap(~State) +
  theme(legend.position = "none")

#now with a geom_text: Not ideal, but this is the best I can do for now

US_ineq %>% 
  ggplot() +
  geom_line(data = transform(US_ineq, State = NULL), aes(Year, Gini), colour = "Black", alpha = 0.5) +
  geom_text(data = (US_ineq %>% filter(Year == max(Year)) %>% 
                    transform(State = NULL)), aes(Year, Gini), label = "US", size = 2) +
  geom_line(data = (inequality %>% filter(State == head(State, n = 15))), aes(Year, Gini, colour = State)) +
  facet_wrap(~State)

```

# Descriptive Stats Income

```{r}

pc_joined %>% 
  filter(State == c("Alabama", "District of Columbia")) %>% 
  ggplot() +
  geom_line(aes(Year, pc_income, colour = State))
  
```


####### Complete Dataset ########

I want dataset with real per capita income and income growth, State, Year, Gini and Gini growth, as well as Regional Groupings

```{r}

joined <- left_join(US_ineq, pcdeflated, by = c("State" = "State", "Year" = "Year")) %>% 
  filter(Measure == "Gini") %>% 
  rename(Gini = Value) %>% 
  select(-Measure)

write.csv(joined, "C:/Projects/Time_Series/PVAR_19491050/data/joined.csv")

```





